String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-09 2023-1-19/sections/recl.tex
revise syntax in recl

%\subsubsection{Syntax}
In the sequel, we define string constraints with regex-counting and string-length function (abbreviated as RECL). The syntax of RECL is defined by the rules in Figure~\ref{fig:syntax}, where $x$ is a string variable, $\anivar$ is an integer variable, $a$ is a character from an alphabet $\Sigma$, and $m, n$ are integer constants. 
A RECL formula $\varphi$ is a conjunction of atomic formulas of the form $x \in \regex$, $x \not \in \regex$, or $t_1\ \op\ t_2$, where $\regex$ is a regular expression,  
$t_1$ and $t_2$ are integer terms, and $\op \in \{=, \neq, \le, \ge, <, >\}$. Atomic formulas of the form $x \in \regex$ or $x \not \in \regex$ are called \emph{regular membership} constraints and atomic formulas of the form $t_1\ \op\ t_2$ are called \emph{length} constraints. 
%
A regular expression $\regex$ is built from $\emptyset$, the empty string $\epsilon$, and characters $a$ by using concatenation  $\cdot$, union $+$, Kleene star $^*$, intersection $\cap$, complement $\bar{\mbox{ }}$, difference $\setminus$, counting $^{\{m,n\}}$ or $^{\{m,\infty\}}$. An integer term is is built from constants $n$, variables $i$, and string lengths $|x|$, by operators $+$ and $-$.
% 
%is a quantifier-free formula that can be a regular membership $x\in\regex$, a quantifier-free Presburger formula $\alpha_1 \leq \alpha_2$, a negation of a formula, or a disjunction of two formulas. $\alpha$ is an integer term which can be an integer constant $m$, integer variable $i$, the length of the string term $|s|$, the minus of an integer term, and the plus of two integer terms. The regular expression $\regex$ is built on empty string $\epsilon$, constant letter $a\in \Sigma$. The supported regex operations involve concatenation $\cdot$, disjunction $+$, intersection $\times$, closure $*$, complement $C$, and bounded repetition $\{m, n\}$. We use $term(\varphi)$ to denote the set of terms and $strvar(\varphi)$ to denote the string variables occurring in $\varphi$. The \emph{literals} of $\varphi$ are $x\in \regex$ and $\alpha_1 \leq \alpha_2$. $x\in \regex$ is called \emph{regular literals} and $\alpha_1 \leq \alpha_2$ is called \emph{linear literals}.
\begin{figure}[h]
  \centering
%  \begin{tabular}{l r}
$ \begin{array}{l l l r}
    \varphi & ::= & x\in \regex \mid t_1\  \op\ t_2 \mid  \varphi \wedge \varphi                                              & \mbox{formulas}            \\
    \regex &::= & \emptyset \mid \epsilon \mid a \mid \regex\cdot \regex \mid \regex+\regex \mid \regex^* \mid \regex \cap \regex \mid \overline{\regex} \mid \regex \setminus \regex \mid \regex^{\{m,n\}} \mid \regex^{\{m,\infty\}} \mid (e) & \mbox{regexes} \\
    t &::= & n \mid \anivar \mid  |x| \mid t - t \mid t + t                                                                    & \mbox{integer terms}
    \end{array}
  $
%  \end{tabular}
  \caption{Syntax of RECL }\label{fig:syntax}
\end{figure}
Moreover, for $S \subseteq \Sigma$ with $S = \{a_1, \cdots, a_k\}$, we use $S$ as an abbreviation of $a_1 + \cdots + a_k$.

The definition of the semantics of RECL is somehow standard. Nevertheless, for completeness, we include the definition of semantics here.

>>>>

%\subsubsection{Syntax}
In the sequel, we define string constraints with regex-counting and string-length function (abbreviated as RECL). The syntax of RECL is defined by the rules in Figure~\ref{fig:syntax}, where $x$ is a string variable, $\anivar$ is an integer variable, $a$ is a character from an alphabet $\Sigma$, and $m, n$ are integer constants. 
A RECL formula $\varphi$ is a conjunction of atomic formulas of the form $x \in \regex$, $x \not \in \regex$, or $t_1\ \op\ t_2$, where $\regex$ is a regular expression,  
$t_1$ and $t_2$ are integer terms, and $\op \in \{=, \neq, \le, \ge, <, >\}$. Atomic formulas of the form $x \in \regex$ or $x \not \in \regex$ are called \emph{regular membership} constraints and atomic formulas of the form $t_1\ \op\ t_2$ are called \emph{length} constraints. 
%
A regular expression $\regex$ is built from $\emptyset$, the empty string $\epsilon$, and characters $a$ by using concatenation  $\cdot$, union $+$, Kleene star $^*$, intersection $\cap$, complement $\bar{\mbox{ }}$, difference $\setminus$, counting $^{\{m,n\}}$ or $^{\{m,\infty\}}$. An integer term is is \denghang{double is} built from constants $n$, variables $i$, and string lengths $|x|$, by operators $+$ and $-$.
% 
%is a quantifier-free formula that can be a regular membership $x\in\regex$, a quantifier-free Presburger formula $\alpha_1 \leq \alpha_2$, a negation of a formula, or a disjunction of two formulas. $\alpha$ is an integer term which can be an integer constant $m$, integer variable $i$, the length of the string term $|s|$, the minus of an integer term, and the plus of two integer terms. The regular expression $\regex$ is built on empty string $\epsilon$, constant letter $a\in \Sigma$. The supported regex operations involve concatenation $\cdot$, disjunction $+$, intersection $\times$, closure $*$, complement $C$, and bounded repetition $\{m, n\}$. We use $term(\varphi)$ to denote the set of terms and $strvar(\varphi)$ to denote the string variables occurring in $\varphi$. The \emph{literals} of $\varphi$ are $x\in \regex$ and $\alpha_1 \leq \alpha_2$. $x\in \regex$ is called \emph{regular literals} and $\alpha_1 \leq \alpha_2$ is called \emph{linear literals}.
\begin{figure}[h]
  \centering
%  \begin{tabular}{l r}
$ \begin{array}{l l l r}
    \varphi & ::= & x\in \regex \mid t_1\  \op\ t_2 \mid  \varphi \wedge \varphi                                              & \mbox{formulas}            \\
    \regex &::= & \emptyset \mid \epsilon \mid a \mid \regex\cdot \regex \mid \regex+\regex \mid \regex^* \mid \regex \cap \regex \mid \overline{\regex} \mid \regex \setminus \regex \mid \regex^{\{m,n\}} \mid \regex^{\{m,\infty\}} \mid (e) & \mbox{regexes} \\
    t &::= & n \mid \anivar \mid  |x| \mid t - t \mid t + t                                                                    & \mbox{integer terms}
    \end{array}
  $
%  \end{tabular}
  \caption{Syntax of RECL }\label{fig:syntax}
\end{figure}
Moreover, for $S \subseteq \Sigma$ with $S = \{a_1, \cdots, a_k\}$, we use $S$ as an abbreviation of $a_1 + \cdots + a_k$.

The definition of the semantics of RECL is somehow\denghang{inappropriate use of somehow} standard. Nevertheless, for completeness, we include the definition of semantics here.
